### 2023-08-29 manuscript.tex
Moved Materials and methods after introduction
Moved figures to figure section
Added acknowledgements and funding text and removed article summary
Added clearpage commands to space references, figures, and tables properly
Removed tables section because tables will remain in supplement
Removed footnote in introduction

\section{Article summary}
insert article summary

\section{Introduction}
Comparative genomics is a powerful tool for yielding insights into evolutionary relationships, molecular function, and the forces that drive gene, genome, and population evolution. These methods often rely on the identification of homologous sequences or homologs, that is sequences with common ancestry, since this ensures that differences between sequences reflect variations in evolution from a common point of divergence. However, many analyses impose the additional condition that the sequences have diverged through speciation events (orthology) rather than duplications (paralogy) or other mechanisms such as horizontal gene transfer~\cite{Fitch1970}. The underlying assumption is orthologs have conserved equivalent functions whereas paralogs, by virtue of their redundancy, are more likely to diverge~\cite{Ohno1970, Nowak1997, Altenhoff2012, Pegueroles2013, Soria2014}.\footnote{For multiple sequences, orthology is usually defined relative to their most recent common ancestor. This technically includes sequences which split by duplication after this point (in-paralogs), but excludes sequences which split by duplication before (out-paralogs) \cite{Remm2001}. Many current orthology inference pipelines explicitly incorporate steps to detect in-paralogs. However, the resulting orthologs groups can easily be restricted to those without in-paralogs (single copy orthologs) for analyses where an assumption of conserved function is necessary.} This relationship between orthology and function is an essential component of modern biological research since it permits the transfer of annotations between biological systems using sequence similarity alone.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{graphical_overview/out/graphical_overview.png}
\centering
\caption{\textbf{Overview of orthology inference pipeline.}
\textbf{(A-C)} The large, open circles represent the annotated genomes, and the small, filled circles represent the protein sequences associated with each annotation. Sequences that share homology are colored with shades of the same hue. \textbf{(D-F)} The small, filled circles and lines represent the same sequences and best hits from the previous steps.}
\label{fig:graphical_overview}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{percolation/out/percolation.png}
\includegraphics[width=\textwidth]{percolation/out/005F.png}
\centering
\caption{\textbf{Clique percolation algorithm.}
\textbf{(A)} Cliques for \textit{k} equal to three, four, five, and six. \textit{k} equal to one and two correspond to a single node and two nodes joined by an edge, respectively. \textbf{(B)} Illustration of clique percolation algorithm where \textit{k = 4}. \textbf{(C)} A single component clustered by clique percolation with varying values of \textit{k}. Nodes are colored according to their cluster. If a node belongs to multiple clusters, it uses a blend of those colors.}
\label{fig:percolation}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hmm/out/hmm_architecture.png}
\includegraphics[width=\textwidth]{hmm/out/2252.png}
\centering
\caption{\textbf{HMM emission architecture and a decoded alignment.}
\textbf{(A)} Schematic of theoretical alignment generating process and corresponding probabilistic components in HMM. White and colored boxes indicate gap and non-gap symbols in the biological sequences, respectively. White and grey boxes indicate gap and non-gap symbols in the encoded sequences, respectively. c\textsubscript{0} and c\textsubscript{1} indicate the first and second columns in the alignment, respectively. The parameters associated with each component are shown to the right. \textbf{(B)} The alignment of the representative sequences in orthologous group 2252 decoded using the trained HMM.}
\label{fig:hmm}
\end{figure}

@@

\section{Acknowledgements}
insert acknowledgements

\section{Funding}
insert funding

@@

\section{Figures}
insert figures

\section{Tables}
insert tables

>>>>

\section{Introduction}
Comparative genomics is a powerful tool for yielding insights into evolutionary relationships, molecular function, and the forces that drive gene, genome, and population evolution. These methods often rely on the identification of homologous sequences or homologs, that is sequences with common ancestry, since this ensures that differences between sequences reflect variations in evolution from a common point of divergence. However, many analyses impose the additional condition that the sequences have diverged through speciation events (orthology) rather than duplications (paralogy) or other mechanisms such as horizontal gene transfer~\cite{Fitch1970}. The underlying assumption is orthologs have conserved equivalent functions whereas paralogs, by virtue of their redundancy, are more likely to diverge~\cite{Ohno1970, Nowak1997, Altenhoff2012, Pegueroles2013, Soria2014}. This relationship between orthology and function is an essential component of modern biological research since it permits the transfer of annotations between biological systems using sequence similarity alone.

@@

@@

\section{Acknowledgements}
We thank the past and present members of the Eisen lab for helpful discussions and their assistance in revising the manuscript.

\section{Funding}
This work was supported by the Howard Hughes Medical Institute.

@@

\clearpage
\printbibliography

\clearpage
\section{Figures}
\begin{figure}[h!]
\includegraphics[width=\textwidth]{graphical_overview/out/graphical_overview.png}
\centering
\caption{\textbf{Overview of orthology inference pipeline.}
\textbf{(A-C)} The large, open circles represent the annotated genomes, and the small, filled circles represent the protein sequences associated with each annotation. Sequences that share homology are colored with shades of the same hue. \textbf{(D-F)} The small, filled circles and lines represent the same sequences and best hits from the previous steps.}
\label{fig:graphical_overview}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{percolation/out/005F.png}
\centering
\caption{\textbf{Clique percolation algorithm.}
\textbf{(A)} Cliques for \textit{k} equal to three, four, five, and six. \textit{k} equal to one and two correspond to a single node and two nodes joined by an edge, respectively. \textbf{(B)} Illustration of clique percolation algorithm where \textit{k = 4}. \textbf{(C)} A single component clustered by clique percolation with varying values of \textit{k}. Nodes are colored according to their cluster. If a node belongs to multiple clusters, it uses a blend of those colors.}
\label{fig:percolation}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hmm/out/2252.png}
\centering
\caption{\textbf{HMM emission architecture and a decoded alignment.}
\textbf{(A)} Schematic of theoretical alignment generating process and corresponding probabilistic components in HMM. White and colored boxes indicate gap and non-gap symbols in the biological sequences, respectively. White and grey boxes indicate gap and non-gap symbols in the encoded sequences, respectively. c\textsubscript{0} and c\textsubscript{1} indicate the first and second columns in the alignment, respectively. The parameters associated with each component are shown to the right. \textbf{(B)} The alignment of the representative sequences in orthologous group 2252 decoded using the trained HMM.}
\label{fig:hmm}
\end{figure}
